Debugging by \textit{lastChange}

### 2010-10-29 papers/pldi2011/sigplanconf-template.tex

Imagine a program execution paused on a breakpoint and a developer
suspicious about the value of an object property or a variable. The
developer queries \textit{lastChange} on the value. The debugger
locates and shows the last change of the object property or
variable. Debugger reproduces the buggy execution and collects data
and once the execution reaches the same place (i.e., the same
breakpoint hit), it pauses the execution, analyzes the collected data
and shows the location of the last change to the developer. The
developer can also examine the program state at the located point of
execution, and continue debugging by more \textit{lastChange} queries
from that point.

We refer to the breakpoint used in \textit{lastChange} as the
\textit{reproduction point}: it is then point the developer knows
reproduces the bug and it is the point were the \textit{lastChange}
algorithm ends. The ability to identify this reproduction point, that
is a test case and a breakpoint after the bug is appears, is the only
prerequisite for \textit{lastChange} functionality.  This means
\textit{lastChange} can be implemented on top of a conventional
breakpoint debugger and it does not require a special environment to
create identical, instruction by instruction, re-executions.

@@

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented it in an extension to the Firebug
JavaScript debugger\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface [] and %\cite{JSD}
\textit{Querypoint} is developed over this interface. Our prototype implementation defines our four primitive operations in JavaScript code and using techniques which are cumbersome and comparatively slow to execute.  A professionally useful debugger would implement these primitive operations in the Javasript engine.

\subsection{\texttt{objectId()} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. To set
the property, we call the \texttt{defineProperty} standard JavaScript
function. \texttt{defineProperty} receives an argument which specifies
whether the property is enumerable or not. By setting the enumerable
field \texttt{false} for \texttt{\_objectId}, this property will not
apear in \texttt{for(property in object)} loops and therfore has no
effect on the program execution. Note that \texttt{\_objectId} is
not set for all objects but only those objects that need an id.

@@

\textit{lastChange} functionality supports obtaining information about
the execution state logically earlier in the control flow. This
support resembles a mixture of replay-based and logging-based
debugging. Replay-based approaches capture limited data during
execution and replay the bug-gy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} functionality collects data on re-execution
but this data is limited to the current queries of developer.

Among replay-based debuggers we compare to bdb \cite{Boothe} and
reverse watchpoint \cite{Maruyama}.  A bidirectional C debugger, bdb
employs a step counter to locate the requested point from the
beginning of execution. It relies on deterministic execution replay
(i.e., the same sequence of instructions in re-execution) and records
the results of non-deterministic system calls and re-injects them into
the program when it is replayed. It makes use of checkpoints to reduce
the time needed for re-execution.  Reverse watchpoint, is proposed by
Maruyama et al., analyses the execution and moves the debugger to the
last write access of a selected variable by re-executing the program
from the beginning\cite{Maruyama}.  Similar to bdb it relies on deterministic
replay and uses a counter to correctly locate a point in the next
execution. The main disadvantage of these approaches is requiring the
exactly the same executions. Even one instruction difference between
two executions leads to wrong results. This requirement is that
much restricting that after ten years from bdb paper publication,
there is no newer implementation of this idea. On the other hand,
\textit{lastChange} doesn't require any special feature in the
re-execution and it is fit to  developers' everyday debugging
practice.

@@

Two new directions in logging debuggers explore more detailed use of
the log and more effective logging approaches. WhyLine\cite{Ko}
provides visual interface to collected runtime information and let
developer to move on execution log using queries expressed in terms of
the programming objects. WhyLine stores the program user in-terface in
addition to program trace and provides answers to why and why not
questions to the user. Jive\cite{Czyz} depicts the history of
execution by a sequence diagram and lets user to query on events
database. Both tools suffer from similar issues with omniscient
debuggers.
%both provide models for extending Querypoint debugging to obtain a better user interface while retaining the flexible conven-tional replay model of debugging.

@@

@@

>>>>

Imagine a program execution paused on a breakpoint and a developer
suspicious about the value of an object property or a variable. The
developer queries \textit{lastChange} on the value. The debugger
reproduces the buggy execution and collects data
and once the execution reaches the same place (i.e., the same
breakpoint hit), it pauses the execution, analyzes the collected data
and shows the location of the last change to the developer. The
developer can also examine the program state at the located point of
execution, and continue debugging by more \textit{lastChange} queries
from that point.

We refer to the breakpoint used in \textit{lastChange} as the
\textit{reproduction point}: this is a point where the developer knows
reproduces the bug and it is the point were the \textit{lastChange}
algorithm ends. The ability to identify this reproduction point, i.e.
 a test case and a breakpoint after the bug is appears, is the only
prerequisite for \textit{lastChange} functionality.  This means
\textit{lastChange} can be implemented on top of a conventional
breakpoint debugger and it does not require a special environment to
create identical, instruction by instruction, re-executions.

@@

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented it in an extension to the Firebug
JavaScript debugger\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface \cite{JSD} and 
\textit{Querypoint} is developed over this interface. Our prototype implementation defines our four primitive operations in JavaScript code and using techniques which are cumbersome and comparatively slow to execute.  A professionally useful debugger would implement these primitive operations in the Javasript engine.

\subsection{\texttt{objectId()} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. To assign the id to the object we 
set the id as a regular object property. This might change
the program behaviour due to the appearance of this property in 
\texttt{for(property in object)} loops. \texttt{defineProperty} is a standard JavaScript
function which is available in the next version of Firefox\footnote[1]{https://developer.mozilla.org/en/JavaScript/Reference/Global\_Objects\\/Object/defineProperty}. \texttt{defineProperty} receives 
an argument which specifies whether the property is enumerable or not. By setting the enumerable
field \texttt{false} for \texttt{\_objectId}, this property will not
apear in \texttt{for(property in object)} loops and therfore has no
effect on the program execution. Note that \texttt{\_objectId} is
not set for all objects but only those objects that need an id.

@@

\textit{lastChange} functionality supports obtaining information about
the execution state logically earlier in the control flow. This
support resembles a mixture of replay-based and logging-based
debugging. Replay-based approaches capture limited data during
execution and replay the buggy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} functionality collects data on re-execution
but this data is limited to the current queries of developer.

Among replay-based debuggers we compare to bdb \cite{Boothe} and
reverse watchpoint \cite{Maruyama}.  A bidirectional C debugger, bdb
employs a step counter to locate the requested point from the
beginning of execution. It relies on deterministic execution replay
(i.e., the same sequence of instructions in re-execution) and records
the results of non-deterministic system calls and re-injects them into
the program when it is replayed. It makes use of checkpoints to reduce
the time needed for re-execution.  Reverse watchpoint, is proposed by
Maruyama et al., analyses the execution and moves the debugger to the
last write access of a selected variable by re-executing the program
from the beginning\cite{Maruyama}.  Similar to bdb it relies on deterministic
replay and uses a counter to correctly locate a point in the next
execution. The main disadvantage of these approaches is requiring the
exactly the same executions. Even one instruction difference between
two executions leads to wrong results. On the other hand,
\textit{lastChange} doesn't require any special feature in the
re-execution and it is fit to  developers' everyday debugging
practice.

@@

@@

Two new directions in logging debuggers explore more detailed use of
the log and more effective logging approaches. WhyLine\cite{Ko}
provides visual interface to collected runtime information and let
developer to move on execution log using queries expressed in terms of
the programming objects. WhyLine stores the program user interface in
addition to program trace and provides answers to why and why not
questions to the user. Jive\cite{Czyz} depicts the history of
execution by a sequence diagram and lets user to query on events
database. Both tools suffer from the requirement of gathering tracing information before their unique capabilities can be used.
We imagine that the runtime queries we use in \textit{lastChange}  may be used to gather data incrementally for these kinds of debugging approaches.

@@

\bibitem[JSD(2010)]{JSD}
Mozila JavaScript Debugging Interface(JSD). \newblock http://www.mozilla.org/js/jsd.
